---
title: "AgingStudy_analysisplan"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(lme4) # For doing linear mixed modeling
library(jtools) # For looking at lmer outputs
library(lmerTest) # For looking at lmer outputs
## Load data files
# Summary data for both tasks
df = read.csv("./Summary.csv", fileEncoding="UTF-8-BOM") 
# Data for individual reaction times

yadf = read.csv("./YA_rawRT.csv") %>%
  filter(Subject != '5c97c01445658c0016752431'& 
         Subject != '60de5412b084f48d5d180a13'& 
         Subject != '60d754d900f33c946d172323'&
         Subject != '60e228622e17c4ed597d92eb'& 
         Subject != '60214feec65df50cd6837a48'& 
         Subject != '5eb4841544c2df2f0190e8a0'& 
         Subject != '60cf8e9061eea6080907595a'& 
         Subject != '606192adc1fba49d7198df8f') 

oadf = read.csv("./OA_rawRT.csv") %>%
  filter(Subject != '5c5dadc66467ac0001b7fa53' & 
           Subject != '5f564265f9d134032e757250'& 
           Subject != '60a8bc1ed3a5087a7538ba3d'&
           Subject != '5780d9a1900cc80001d2d1c2'&
           Subject != '5a69d956111b250001e342b3'&
           Subject != '5ea02e5c3fb6e6000a30693b'&
           Subject != '5f607efeea3f070107273203'& 
           Subject != '5b100cb0a3553300017541c8'&
           Subject != '5ea00e67848a4300085fa5cd'&
           Subject != '5f6f7258aa7332501adb94f4'&
           Subject != '60c0c4333e49151d768bb35b')
oadf = mutate(oadf, Age = 0 )
yadf = mutate(yadf, Age = 1)
rtdf = rbind(oadf, yadf)

rtdf$RT = as.numeric(rtdf$RT*1000)
```

```{r Filter loaded dataset}
# Change here to change filtering criteria
df = df %>%
  filter(Exclusion != 'exc') # remove the excluded participatns
  #filter(Exclusion != 'out') # uncomment when removing outliers 

# Contrast code Age groups
df$Age = as.factor(df$Age)
df = mutate(df, coded_Age = Age)
contrasts(df$coded_Age) = contr.treatment(2)

df$Gender = as.factor(df$Gender)
df = mutate(df, coded_Gender = Gender)
contrasts(df$coded_Gender) = contr.treatment(2)

df$Counterbalance = as.factor(df$Counterbalance)
df = mutate(df, coded_Counterbalance = Counterbalance)
contrasts(df$coded_Counterbalance) = contr.treatment(2)

```

### Hypothesis
Pattern separation and statistical learning are two independent processes. Aging, which typically impaires the dentate gyrus and pattern separation, does not affect statistical learning which does not depend on the dentate gyrus.

**Prediction**: Older adults will be significantly worse at *lure discrimination* (but not recognition) on the auditory MST task compared to younger adults, but will perform just as well as the younger adults on the statistical learning task.

### Sample size calculation (GPower)

  
	F tests - ANOVA: Fixed effects, omnibus, one-way
	Analysis: A priori: Compute required sample size
	Input: Effect size f = 0.4200840
	α err prob = 0.05
	Power (1-β err prob) = 0.80
	Number of groups = 2
	Output:	Noncentrality parameter λ = 8.4705872
	Critical F = 4.0517487
	Numerator df = 1
	Denominator df = 46
	Total sample size	=	48
	Actual power = 0.8130553
	
**= 24 + 4 subjects per age group** to achieve 80% power to detect a large effect size (f = 0.42, from the literature)

### Primary analysis
Variables of interest: 
- Age group (1-4)
- Lure discrimination (1)
- Familiarity rating (2)
- AFC accuracy (3)
- Reaction time (4)

Control variables (included in the full model):
- Gender
- Counterbalance condition
- Randomization condition (both MST and SL)
- Occurrence (for RT)

1. Group difference in lure discrimination
```{r Lure discrimination difference - GLM}
# Run GLM
discrimination_full = lm(Disc ~ coded_Age + coded_Gender + coded_Counterbalance, data = df)
discrimination_reduced = lm(Disc ~ coded_Age, data = df)
summary(discrimination_reduced)
summary(discrimination_full)
anova(discrimination_full, discrimination_reduced)

disc_anova = aov(Disc ~ coded_Age, data = df)
summary(disc_anova)
```

2. Group difference in Familiarity rating
```{r Familiarity difference - GLM}
df = mutate(df, fam_rating_diff = Word - Nonword)
longer_df = pivot_longer(df, cols=ends_with('ord'), names_to = 'word_type', values_to = 'ratings')
longer_df$word_type = as.factor(longer_df$word_type)
contrasts(longer_df$word_type) = contr.treatment(3, base=1)
# Run linear mixed model with subject as random effects
familiarity_full = lmer(ratings ~ word_type*coded_Age + coded_Gender + coded_Counterbalance + (1|Subject), 
                        data = longer_df)
familiarity_model = lmer(ratings ~ word_type + coded_Age + (1|Subject), data = longer_df)
summary(familiarity_full)
```

3. Group difference in AFC accuracy
```{r AFC difference - GLM}
# Run GLM
AFC_full = lm(AFC ~ coded_Age + coded_Gender + coded_Counterbalance, data = df)
AFC_reduced = lm(AFC ~ coded_Age, data = df)
summary(AFC_full)
summary(AFC_reduced)
anova(AFC_full, AFC_reduced)
```

4. Group difference in RT effect
```{r RT effect difference - GLM}
#rtdf = rtdf %>%
#  filter(Subject != '60dc4b08dd9ec36f5ca9e3fa')

rtdf[, 'Coded_Position'] = as.factor(rtdf[, 'Position'])
rtdf[, 'Coded_Occurrence'] = as.factor(rtdf[, 'Occurrence'])
rtdf[, 'Coded_Condition'] = as.factor(rtdf[, 'Condition'])
rtdf[, 'Age'] = as.factor(rtdf[, 'Age'])
contrasts(rtdf$Coded_Position) = contr.treatment(3)
contrasts(rtdf$Coded_Occurrence) = contr.treatment(3)
contrasts(rtdf$Coded_Condition) = contr.treatment(3)
contrasts(rtdf$Age) = contr.treatment(2)

position_full = lmer(RT ~ Coded_Position*Coded_Occurrence + Coded_Condition + Age*Coded_Position + (1 + Coded_Position| Subject), data = rtdf)
position_reduced = lmer(RT ~  Coded_Position + (1 + Coded_Position | Subject), data=rtdf)
summary(position_full)
summary(position_reduced)
#summ(positionEffect)
anova(position_full, position_reduced)

```
### Secondary analysis

1. Individual RT effect
```{r Individual RT effect}
subject_list = as.list(distinct(rtdf, Subject))
pvalue_list = data.frame(Subject = numeric(0), pval = numeric(0))
for (i in subject_list){
  for (x in i){
    single_subj = rtdf %>%
      filter(Subject == x)
    positionEffect = lm(RT ~ Coded_Position, data = single_subj)
    print(x)
    print(summary(positionEffect))
    pvalue = summary(positionEffect)$coefficients[3, 4]
    #pvalue_list[[length(pvalue_list)+1]] = c(x, pvalue)
    #pvalue_list = rbind(pvalue_list, c(x, pvalue))
    pvalue_list[nrow(pvalue_list)+1, ] = c(x, pvalue)
  }
}

# Output p-value for the difference between 1st and 3rd positions for each individual
#write.table(pvalue_list, "./YA_rawRT.csv", append=TRUE, sep = ',')
```

2. Group difference in recognition
```{r Recognition difference}
# Run GLM
recognition_full = lm(Recog ~ coded_Age + coded_Gender + coded_Counterbalance, data = df)
recognition_reduced = lm(Recog ~ coded_Age, data = df)
summary(recognition_full)
summary(recognition_reduced)
anova(recognition_reduced, recognition_full)

recog_anova = aov(Recog ~ coded_Age, data = df)
summary(recog_anova)
```

3. SL-MST correlation
```{r SL-MST Correlation}
library(Hmisc)

correl_df = df %>%
  mutate(RT.Diff = Diff) %>%
  select(Recog, Disc, Fam.Diff, AFC, RT.Diff)

correl_res = cor(correl_df)

correl_res_p = rcorr(as.matrix(correl_df))
correl_res_p

```


#### Visualization
1. Position effect
```{r Visualize position-wise}
rtdf$Position = as.factor(rtdf$Position)

confidence_interval = function(x){ # Function that calculates the 95% confidence interval of distribution x
  qnorm(.95) * (sd(x)/sqrt(length(x)))
}

# Summarize based on syllable positions
sum_pos = rtdf %>%
  dplyr::group_by(Position) %>%
  dplyr::summarise(mean = mean(RT), CI = confidence_interval(RT))

# Plot
position_plot = ggplot(sum_pos, aes(x=Position, y=mean)) +
  #geom_bar(stat = "identity", fill = 'black', alpha = 0.5, width = 0.5) + #For bar graph
  geom_line(aes(group = 1)) +
  geom_point() +
  ylab('Mean RT (ms)') +
  xlab('Syllable position') +
  geom_errorbar(aes(ymin = mean-CI, ymax = mean+CI, width=0.05)) +
  theme(axis.line = element_line(color = 'black')) +
## Uncomment for a blank background
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"))

position_plot
```

2. Syllable effect
```{r Visualize syllable-wise}
rtdf$Syllable = as.factor(rtdf$Syllable)

confidence_interval = function(x){ # Function that calculates the 95% confidence interval of distribution x
  qnorm(.95) * (sd(x)/sqrt(length(x)))
}

# Summarize based on syllables
sum_syl = rtdf %>%
  dplyr::group_by(Syllable) %>%
  dplyr::summarise(mean = mean(RT), CI = confidence_interval(RT))

# Plot
syllable_plot = ggplot(sum_syl, aes(x=Syllable, y=mean)) +
  geom_bar(stat = "identity", fill = 'black', alpha = 0.5, width = 0.5) + #For bar graph
  #geom_line(aes(group = 1)) +
  geom_point() +
  ylab('Mean RT (ms)') +
  xlab('Syllable position') +
  coord_cartesian(ylim=c(200, 630)) +
  geom_errorbar(aes(ymin = mean-CI, ymax = mean+CI, width=0.05)) +
  theme(axis.line = element_line(color = 'black'))
## Uncomment for a blank background
  #theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"))

syllable_plot

summary(rtdf)
```

3. Occurrence x Position interaction
```{r Visualize Occurrence x Position}
rtdf$Occurrence = as.factor(rtdf$Occurrence)
rtdf$Position = as.numeric(rtdf$Position)

confidence_interval = function(x){ # Function that calculates the 95% confidence interval of distribution x
  qnorm(.95) * (sd(x)/sqrt(length(x)))
}

# Summarize based on syllable positions and occurrences
#positions = c(sum_OP$mean_1, sum_OP$mean_2, sum_OP$mean_3)

sum_PO = rtdf %>%
  dplyr::group_by(Position, Occurrence) %>%
  dplyr::summarise(mean = mean(RT), CI = confidence_interval(RT)) %>%
  pivot_wider(names_from = 'Occurrence', values_from = c('mean', 'CI'))

# Plot
ggplot(sum_PO, aes(x=Position)) +
  #geom_bar(stat = "identity", fill = 'black', alpha = 0.5, width = 0.5) + #For bar graph
  geom_line(aes(y=mean_1), color = 'blue') +
  geom_line(aes(y=mean_2), color = 'green') +
  geom_line(aes(y=mean_3), color = 'red') +
  ylab('Mean RT (ms)') +
  xlab('Syllable position') +
  geom_errorbar(aes(ymin = mean_1-CI_1, ymax = mean_1+CI_1, width=0.05)) +
  geom_errorbar(aes(ymin = mean_2-CI_2, ymax = mean_2+CI_2, width=0.05)) +
  geom_errorbar(aes(ymin = mean_3-CI_3, ymax = mean_3+CI_3, width=0.05)) +
  theme(axis.line = element_line(color = 'black')) +
## Uncomment for a blank background
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"))

#OP_plot

```

4. d-prime distributions
```{r Boxplot d-prime for Disc and Recog}
plot_disc = ggplot(df, aes(x = Age, y = Disc), narm = TRUE) +
  geom_boxplot(fill='light gray') +
  geom_point() +
  ylab("Discrimination") +
  xlab("") +
  theme(axis.line = element_line(color = 'black')) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"))

plot_recog = ggplot(df, aes(x = Age, y = Recog), narm = TRUE) +
  geom_boxplot(fill='light gray') +
  geom_point() +
  ylab("Recognition") +
  xlab("") +
  theme(axis.line = element_line(color = 'black')) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"))

plot_disc
plot_recog

```

```{r Visualize correlations}
ggplot(df, aes(x=Disc, y=Fam.Diff)) +
  geom_point() +
  ylab("Familiarity rating (Word - Nonword)") +
  xlab("Lure Discrimination") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"))

plot(df$Disc, df$Fam.Diff)
lines(predict(lm(df$Disc~df$Fam.Diff)),col='black')

ggplot(df, aes(x=Disc, y=AFC)) +
  geom_point() +
  ylab("AFC Accuracy") +
  xlab("Lure Discrimination") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"))

ggplot(df, aes(x=Fam.Diff, y=AFC)) +
  geom_point() +
  xlab("Familiarity rating (Word - Nonword)") +
  ylab("AFC Accuracy") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"))
```


